A Comparison of Dimensionality Reduction Techniques for Robotic Motion Planning

### 2016-05-14 paper/paper.tex
Finished results section

@@

@@

After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
PCA, and Randomized PCA to reduce the dimensionality of the space by using the
sampled points within each $n$-sphere as training data. We then ran PRM* and
RRTConnect on each of these reduced dimensionality spaces, as well as on the
fully-dimensional space as a control. We then ran an inverse transform on the
paths that we find to return them to the fully-dimensional space in order to
verify that the path in the reduced dimension did not come into collision with
obstacles. We collected data on the number of runs that fail to find a
solution, the time it takes to run the algorithm, and the length of the
solution if one was found.  A summary of this paragraph can be found in
Algo.~\ref{alg:DR}.

@@

\section{Results}

@@

Fig.~\ref{fig:plots} shows plots of various measures of performance using
different methods of dimensionality reduction for motion planing. The left and
right columns show the results for RRTConnect and PRM algorithms respectively.
The first, second, and third row show the average path length, computational
time in seconds, and number of failed runs out of 100 respectively. For each
dimensionality reduction technique, there are three bars presenting the results
for 250, 300, and 350 randomly generated obstacles respectively. From the data
we can see that both planners produced similarly sized paths for each
dimensionality reduction technique, and that by not reducing the dimension, the
planner is able to find a path roughly two times shorter.

@@

The RRTConnect algorithm had a much lower failure rate than PRM in all
respective dimensionality reduction technique. This is due to the fact that
RRTConnect is growing a tree towards the goal and PRM is randomly sampling and
connecting points to a growing graph. For both algorithms, a similar trend
exists for failure rate for planning in the reduced dimension space.

\begin{algorithm}[ht] 
    \caption{$\Function{PRM}$~\cite{karaman2011sampling}}
    \label{alg:PRM}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}

\begin{algorithm}[ht] 
    \caption{$\Function{Extend\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
    \label{alg:extend_RRT}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \STATE $ q_{near} \leftarrow \Function{NEAREST\_NEIGHBOR($q$,$\;\mathcal{T}$)} $
        \IF{ $ \Function{NEW\_CONFIG($q$,$\;q_{near}$,$\;q_{new}$)} $}
            \STATE $ \mathcal{T}.\text{add\_vertex($q_{new}$)}; $
            \STATE $ \mathcal{T}.\text{add\_edge($q_{near}$,$\;q_{new}$)}; $
            \IF{$ q_{new} = q $}
                \RETURN $Reached$;
            \ELSE
                \RETURN $Advanced$;
            \ENDIF
        \ENDIF
        \RETURN $Trapped$;
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[ht] 
    \caption{$\Function{Connect\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
    \label{alg:connect_RRT}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \REPEAT 
            \STATE $S \leftarrow \Function{EXTEND\_RRT($\mathcal{T}$,$\;q$)};$ 
        \UNTIL {$ \textbf{not} \; (S = Advanced)$}
        \RETURN $S;$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[ht] 
    \caption{$\Function{RRTConnect($q_{init}$,$\;q_{goal}$)}$~\cite{kuffner2000rrt}}
    \label{alg:RRTConnect}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \STATE $\mathcal{T}_a$.init($q_{init}$); $ \mathcal{T}_b$.init($q_{goal}$);
        \FOR{$k = 1 \ \TO \ K$}
            \STATE $q_{rand} \leftarrow \Function{RANDOM\_CONFIG()}$;
            \IF {$\textbf{not} \ (\Function{EXTEND\_RRT($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
                \IF {$ (\Function{CONNECT\_RRT($\mathcal{T}_b,q_{new}$}) = Reached $}
                    \RETURN $ \Function{PATH($\mathcal{T}_a,\mathcal{T}_b$)} $
                \ENDIF
            \STATE \Function{SWAP($\mathcal{T}_a,\mathcal{T}_b)$}
            \ENDIF
        \ENDFOR
        \RETURN $Failure$
    \end{algorithmic}
\end{algorithm}

% \begin{algorithm}[ht] 
%     \caption{$\Function{RRT}^*$~\cite{karaman2011sampling}}
%     \label{alg:RRTConnect}
%     \begin{algorithmic}[1]
%         \setcounter{ALC@line}{0}
%         \vspace*{1mm}

%         \STATE $V \leftarrow \{x_{\Var{init}}\}$
%         \STATE $E \leftarrow \emptyset$
%         \FOR{$i = 1 \ \TO \ n$}
%             \STATE $\xr \leftarrow \Var{SampleFree}_i$
%             \STATE $\xnearest \leftarrow \Function{Nearest}(G = (V, E), \xr)$
%             \STATE $\xnew \leftarrow \Function{Steer}(\xnearest, \xr)$
%             \IF{$\Function{ObstacleFree}(\xnearest, \xnew)$}
%                 \STATE $\xnear \leftarrow \Function{Near}(G = (V, E), \xnew,
%                 \min \{\gamma \cdot \Function{RRT}(\log(|V|) / |V|)^{1 / d}, \eta\})$
%                 \STATE $V \leftarrow V \cup \{\xnew\}$
%                 \STATE $\xmin \leftarrow \xnearest$
%                 \STATE $\cmin \leftarrow \Function{Cost}(\xnear) +
%                 c(\Function{Line}(\xnearest, \xnew))$
%                 \FORALL{$\xnear \in X_{\Var{near}}$}
%                     \STATE $C \leftarrow \cost(\xnear) + \linef{\xnear, \xnew}$
%                     \IF{$\Function{CollisionFree}(\xnear, \xnew) \wedge C < \cmin$}
%                         \STATE $\xmin \leftarrow \xnear$
%                         \STATE $\cmin \leftarrow C$
%                     \ENDIF
%                     \STATE $E \leftarrow E \cup \{(\xmin, \xnew)\}$
%                 \ENDFOR
%                 \FORALL{$\xnear \in X_{\Var{near}}$}
%                     \STATE $C \leftarrow \cost(\xnew) + \linef{\xnew, \xnear}$
%                     \IF{$\Function{CollisionFree}(\xnew, \xnear) \wedge C < \cost(\xnear)$}
%                         \STATE $\xparent \leftarrow \Function{Parent}(\xnear)$
%                         \STATE $E \leftarrow (E \backslash \{(\xparent, \xnear)\}) \cup \{(\xnew, \xnear)\}$
%                     \ENDIF
%                 \ENDFOR
%             \ENDIF
%         \ENDFOR
%         \RETURN $G = (V, E)$

%     \end{algorithmic}
% \end{algorithm}

\begin{algorithm}[ht] 
    \caption{$\Function{Dimensionality Reduction}$}
    \label{alg:DR}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}

>>>>

\begin{algorithm}[h!] 
    \caption{$\Function{PRM}$~\cite{karaman2011sampling}}
    \label{alg:PRM}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}

@@

\begin{algorithm}[h!] 
    \caption{$\Function{Extend\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
    \label{alg:extend_RRT}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \STATE $ q_{near} \leftarrow \Function{NEAREST\_NEIGHBOR($q$,$\;\mathcal{T}$)} $
        \IF{ $ \Function{NEW\_CONFIG($q$,$\;q_{near}$,$\;q_{new}$)} $}
            \STATE $ \mathcal{T}.\text{add\_vertex($q_{new}$)}; $
            \STATE $ \mathcal{T}.\text{add\_edge($q_{near}$,$\;q_{new}$)}; $
            \IF{$ q_{new} = q $}
                \RETURN $Reached$;
            \ELSE
                \RETURN $Advanced$;
            \ENDIF
        \ENDIF
        \RETURN $Trapped$;
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h!] 
    \caption{$\Function{Connect\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
    \label{alg:connect_RRT}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \REPEAT 
            \STATE $S \leftarrow \Function{EXTEND\_RRT($\mathcal{T}$,$\;q$)};$ 
        \UNTIL {$ \textbf{not} \; (S = Advanced)$}
        \RETURN $S;$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[!h] 
    \caption{$\Function{RRTConnect($q_{init}$,$\;q_{goal}$)}$~\cite{kuffner2000rrt}}
    \label{alg:RRTConnect}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}
        \STATE $\mathcal{T}_a$.init($q_{init}$); $ \mathcal{T}_b$.init($q_{goal}$);
        \FOR{$k = 1 \ \TO \ K$}
            \STATE $q_{rand} \leftarrow \Function{RANDOM\_CONFIG()}$;
            \IF {$\textbf{not} \ (\Function{EXTEND\_RRT($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
                \IF {$ (\Function{CONNECT\_RRT($\mathcal{T}_b,q_{new}$}) = Reached $}
                    \RETURN $ \Function{PATH($\mathcal{T}_a,\mathcal{T}_b$)} $
                \ENDIF
            \STATE \Function{SWAP($\mathcal{T}_a,\mathcal{T}_b)$}
            \ENDIF
        \ENDFOR
        \RETURN $Failure$
    \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h!] 
    \caption{$\Function{Dimensionality Reduction}$}
    \label{alg:DR}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}

After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
PCA, and Randomized PCA to reduce the dimensionality of the space by using the
sampled points within each $n$-sphere as training data. We then ran PRM* and
RRTConnect on each of these reduced dimensionality spaces, as well as on the
fully-dimensional space as a control. We then ran an inverse transform on the
paths that we find to return them to the fully-dimensional space in order to
verify that the path in the reduced dimension did not come into collision with
obstacles. We collected data on the number of runs that fail to find a
solution, the time it takes to run the algorithm, and the length of the
solution if one was found. Note, that the length of the path is computed after
it has been transformed back into the original high dimensional space in order
to make a fair comparison. A summary of this paragraph can be found in
Algo.~\ref{alg:DR}.

@@

\section{Results \& Discussion}

Fig.~\ref{fig:plots} shows plots of various measures of performance using
different methods of dimensionality reduction for motion planing. Please note
that the Y-axes of these plots are not equivalent in the left and right column.
The left and right columns show the results for RRTConnect and PRM algorithms
respectively.  The first, second, and third row show the average path length,
computational time in seconds, and number of failed runs out of 100
respectively. For each dimensionality reduction technique, there are three bars
presenting the results for 250, 300, and 350 randomly generated obstacles
respectively. From the data we can see that both planners produced similarly
sized paths for each dimensionality reduction technique, and that by not
reducing the dimension, the planner is able to find a path roughly two times
shorter.

@@

@@

The RRTConnect algorithm had a much lower failure rate than PRM in all
respective dimensionality reduction technique. This is due to the fact that
RRTConnect is growing a tree towards the goal and PRM is randomly sampling and
connecting points to a growing graph. For both algorithms, a similar trend
exists for failure rates when planning in the reduced dimension space. For both
planners, Feature Agglomeration failed more frequently than the rest of the
techniques, and planning in the high dimensional space had the smallest failure
rate. The higher failure rate for Feature Agglomeration is because Feature
Agglomeration works best when features from the data can be combined due to
repetition and redundancy. Since this data was in the $\mathbb{R}^n$ and the
random obstacles were generated all over the space, Feature Agglomeration
failed to produce an embedding that also worked for planning in the lower
dimensional space. Aside from Feature Agglomeration, the dimensionality
reduction techniques failed to find a path to the goal configuration between 20
and 60 percent of the time for PRM and between 5 and 35 percent for RRTConnect.
This is due to the projection of the higher dimensional obstacles into a lower
dimension causing occlusions in the lower dimensional space where they would
not exist in the higher dimension. However, these dimensionality reduction
techniques can be seen as independent random variables for each run, therefore,
you would be able to run the algorithm again with the same probability of
failure. As the number of obstacles increased, so did the failure rate because
the lower dimensional spaces become more densely populated with obstacles
causing more occlusions.

From the data, we can see that dimensionality reduction techniques can reduce
the optimality of the planning algorithm, but in multi-query planning
algorithms, these techniques can significantly reduce the time needed to
compute a solution. Based on the data, it seems that the Truncated SVD
algorithm can be used to produce the lowest number of failures and produces
similar results in the path length and computation time  as compared to the
other techniques used. Also, we Can see that Feature Agglomeration fails to
find a solution path more frequently than the other techniques, but produces
shorter paths when it does find a path.

\section{Conclusion}
